### 2019-06-23 papers/haskell19/1-intro.tex
NIT

% --------------------------------------
% Intro
%
Random property-based testing is a powerful technique for finding bugs
~\cite{ArtsHNS15, HughesPAN16, HughesNSA16, MidtgaardJKNN17}.
%
In Haskell, QuickCheck is the predominant tool for this task \cite{ClaessenH00}.
%
The developers specify
%
\begin{inparaenum}[(i)]
\item the testing properties their systems must fulfill, and
\item random data generators (or generators for short) for the data types
  involved at their properties.
\end{inparaenum}
%
Then, QuickCheck generates random values using these generators, and uses them
to evaluate the testing properties in search of possible counterexamples, which
often indicate the presence of bugs.

@@

This problem, however, can be alleviated by considering additional sources of
structural information from our codebase other than just type definitions
\cite{Mista2019GeneratingRS}.
%
% For instance, having a set of high-level combinators, defining an abstract
% interface of our data types, e.g., |hr :: Html| or |bold :: Html -> Html|; it is
% possible to include them into the generation process, thus obtaining more
% complex and structured values.
%
For instance, it is possible to include into the generation process a set of
high-level combinators of our data types, e.g., |hr :: Html| or |bold :: Html ->
Html|---thus obtaining more complex and structured values.
%
This methodology of generating random data using high-level combinators has
shown to be particulary useful in the presence of monadic code
\cite{Claessen:2002, grieco2017}.
%
% In the same way, having functions defined using pattern matching over the data
% types of our interest, e.g. |f (Join (Text t) (Tag s x)) = dots|, it is also
% possible to include this rich structure which exposes a relationship among
% different data constructors into our random generators.
%
In the same way, it is possible to include into the generation process the
pattern matchings that different functions perform over the data types of our
interest, e.g. |f (Join (Text t) (Tag s x)) = dots|.
%
Such patterns often expose interesting relationships between multiple data
constructors, a valuable asset for testing complex systems expecting highly
structured inputs \cite{klein2009randomized}.
%
% This idea is addressed in depth in the next section, while the rest of this work
% focuses on implementing it in a compositional way.
%
This work shows how to modularly include these extra sources of structural
information into our automatically derived generators in a compositional manner.

%
% Contributions
%
The main contribution of this paper are:
%
\begin{CompactItemize}
\item We identify some important limitations of performing random testing with
  automatically derived generators using state-of-the-art tools (Section
  \ref{sec:sources}).
\item We present an extensible mechanism for representing random values built
  upon different sources of structural information using Data Types \`a la Carte
  (Section \ref{sec:representation}).
\item We develop a modular generation schema, exploiting the composable nature
  of our representations, which are extended to encode information relevant to
  the generation process at the type level (Section \ref{sec:generators}).
\item We propose a simple type-level idiom for describing extensible generators,
  based on the types used to represent the desired shape of our random data
  (Section \ref{sec:typelevel}).\looseness=-1
\item We provide a Template Haskell tool%
  %
  \footnote{Link will be made available for the final version of this paper.}
  %
  for automatically deriving all the required machinery presented throughout
  this paper, and evaluate its generation performance using three real-world
  case studies (Section \ref{sec:casestudies}).
\end{CompactItemize}
%
Overall, we present a novel technique for reusing automatically derived
generators in a composable fashion, in contrast to the usual paradigm of
automatically deriving ``rigid'', monolithic generators.

>>>>

% --------------------------------------
% Intro
%
Random property-based testing is a powerful technique for finding bugs
~\cite{ArtsHNS15, HughesPAN16, HughesNSA16, MidtgaardJKNN17}.
%
In Haskell, QuickCheck is the predominant tool for this task \cite{ClaessenH00}.
%
The developers specify
%
\begin{inparaenum}[(i)]
\item the testing properties their systems must fulfill, and
\item random data generators (or generators for short) for the data types
  involved at their properties.
\end{inparaenum}
%
Then, QuickCheck generates random values using these generators, and uses them
to evaluate the testing properties in search of possible counterexamples, which
always indicate the presence of bugs.

@@

This problem, however, can be alleviated by considering additional sources of
structural information from our codebase other than just type definitions
\cite{Mista2019GeneratingRS}.
%
% For instance, having a set of high-level combinators, defining an abstract
% interface of our data types, e.g., |hr :: Html| or |bold :: Html -> Html|; it is
% possible to include them into the generation process, thus obtaining more
% complex and structured values.
%
For instance, it is possible to include into the generation process a set of
high-level combinators of our data types, e.g., |hr :: Html| or |bold :: Html ->
Html|---thus obtaining more complex and structured values.
%
This methodology of generating random data using high-level combinators has
shown to be particulary useful in the presence of monadic code
\cite{Claessen:2002, grieco2017}.
%
% In the same way, having functions defined using pattern matching over the data
% types of our interest, e.g. |f (Join (Text t) (Tag s x)) = dots|, it is also
% possible to include this rich structure which exposes a relationship among
% different data constructors into our random generators.
%
In the same way, it is possible to include into the generation process the
pattern matchings that different functions perform over the data types of
interest, e.g. |f (Join (Text t) (Tag s x)) = dots|.
%
Such patterns often expose interesting relationships between multiple data
constructors, a valuable asset for testing complex systems expecting highly
structured inputs \cite{klein2009randomized}.
%
% This idea is addressed in depth in the next section, while the rest of this
% work focuses on implementing it in a compositional way.
%
This work shows how to modularly include these extra sources of structural
information into our automatically derived generators in a compositional manner.

%
% Contributions
%
The main contribution of this paper are:
%
\begin{CompactItemize}
\item We identify some important limitations of performing random testing with
  automatically derived generators using state-of-the-art tools (Section
  \ref{sec:sources}).
\item We present an extensible mechanism for representing random values built
  upon different sources of structural information using Data Types \`a la Carte
  (Section \ref{sec:representation}).
\item We develop a modular generation schema, exploiting the composable nature
  of our representations, which are extended to encode information relevant to
  the generation process at the type level (Section \ref{sec:generators}).
\item We propose a simple type-level idiom for describing extensible generators,
  based on the types used to represent the desired shape of our random data
  (Section \ref{sec:typelevel}).\looseness=-1
\item We provide a Template Haskell tool%
  %
  \footnote{Link to be made available in the final version of this paper.}
  %
  for automatically deriving all the required machinery presented throughout
  this paper, and evaluate its generation performance using three real-world
  case studies (Section \ref{sec:casestudies}).
\end{CompactItemize}
%
Overall, we present a novel technique for reusing automatically derived
generators in a composable fashion, in contrast to the usual paradigm of
automatically deriving ``rigid'', monolithic generators.

### 2019-06-23 papers/haskell19/2-sources.tex

In this section, we briefly introduce the motivation for including extra
information into our random generators, and discuss about how this can be
implemented in practice.

@@

A different challenge appears when we try to test functions comprising complex
pattern matchings. Consider for instance the following function for simplifying
|Html| values:

@@

In our tests, we were not able to exercise any of these two patterns more than
6\% of the overall testing time (and considering recursive sub-calls), using
random generators derived using both \emph{MegaDeTH} and \emph{DRAGEN}.
%
As expected, most of the random test cases were exercising the simplest (an
rather uninteresting) patterns of this function, corresponding to its last three
clauses.

@@

% --------------------------------------
% Wrapping up
%
In general, we think that including this static information in our generators
directly is far from ideal, as in many cases we would like to select which
random constructions are of our interest when testing different properties or
sub-systems.
%
In this light, having ``monolithic'' random generators derived automatically
would complicate our desired workflow.
%
We would need to derive different concrete random generators for each use case
instead of having a single |Arbitrary| instance that can be used everywhere.

>>>>

In this section, we briefly introduce the motivation for including extra
information into our random generators, and discuss how this can be implemented
in practice.

@@

A different challenge appears when we try to test functions involving complex
pattern matchings. Consider for instance the following function for simplifying
|Html| values:

@@

In our tests, we were not able to exercise any of these two patterns more than
6\% of the overall testing time (and considering recursive sub-calls), using
random generators derived using both \emph{MegaDeTH} and \emph{DRAGEN}.
%
As expected, most of the random test cases were exercising the simplest (and
rather uninteresting) patterns of this function, corresponding to its last three
clauses.

@@

% --------------------------------------
% Wrapping up
%
In general, we think that including this static information in our generators
directly is far from ideal, as in many cases we would like to select which
random constructions are of interest when testing different properties or
sub-systems.
%
In this light, having ``monolithic'' random generators derived automatically
would complicate our desired workflow.
%
We would need to derive different concrete random generators for each use case
instead of having a single |Arbitrary| instance that can be used everywhere.
